### 2024-06-24 basic-statistics.qmd
20240624 - attempt to fix error
20240624 - more power tests

$$
\begin{aligned}
  f &= \sqrt{\frac{R^2}{1 - R^2}} \\
    &= \sqrt{\frac{\eta^2}{1 - \eta^2}} \\
\end{aligned}
$$ {#eq-cohensF}

@@

$$
\begin{aligned}
  f^2 &= \frac{R^2}{1 - R^2} \\
    &= \frac{\eta^2}{1 - \eta^2} \\
\end{aligned}
$$ {#eq-cohensFsquared}

@@

$$
\begin{aligned}
  R^2 &= 1 - \frac{\sum (Y_i - \hat{Y}_i)^2}{\sum (Y_i - \bar{Y})^2} \\
      &= 1 - \frac{SS_{\text{residual}}}{SS_{\text{total}}} \\
      &= 1 - \frac{\text{sum of squared residuals}}{\text{total sum of squares}}

\end{aligned}
$$ {#eq-rSquared}

@@

```{r}
power <- .8
effectSize_d <- .5
effectSize_r <- .24
```

@@

@@

@@

>>>>

$$
\begin{aligned}
  f &= \sqrt{\frac{R^2}{1 - R^2}} \\
    &= \sqrt{\frac{\eta^2}{1 - \eta^2}}
\end{aligned}
$$ {#eq-cohensF}

@@

$$
\begin{aligned}
  f^2 &= \frac{R^2}{1 - R^2} \\
      &= \frac{\eta^2}{1 - \eta^2}
\end{aligned}
$$ {#eq-cohensFsquared}

@@

$$
\begin{aligned}
  R^2 &= 1 - \frac{\sum (Y_i - \hat{Y}_i)^2}{\sum (Y_i - \bar{Y})^2} \\
      &= 1 - \frac{SS_{\text{residual}}}{SS_{\text{total}}} \\
      &= 1 - \frac{\text{sum of squared residuals}}{\text{total sum of squares}}
\end{aligned}
$$ {#eq-rSquared}

@@

```{r}
power <- .8
effectSize_d <- .5
effectSize_r <- .24
effectSize_f <- .25
effectSize_fSquared <- .06
```

@@

```{r}
pwr::pwr.anova.test(
  n = N,
  f = effectSize_f,
  sig.level = alpha,
  k = numGroups)
```

```{r}
pwr::pwr.anova.test(
  f = effectSize_f,
  power = power,
  sig.level = alpha,
  k = numGroups)
```

```{r}
pwr::pwr.anova.test(
  n = N,
  power = power,
  sig.level = alpha,
  k = numGroups)
```

@@

```{r}
pwr::pwr.r.test(
  n = N,
  r = effectSize_r,
  sig.level = alpha,
  alternative = "two.sided")
```

```{r}
pwr::pwr.r.test(
  r = effectSize_r,
  power = power,
  sig.level = alpha,
  alternative = "two.sided")
```

```{r}
pwr::pwr.r.test(
  n = N,
  power = power,
  sig.level = alpha,
  alternative = "two.sided")
```

@@

```{r}
pwr::pwr.f2.test(
  f2 = effectSize_fSquared,
  sig.level = alpha,
  u = numPredictors,
  v = N - numPredictors - 1)
```

Solving for sample size per group needed (given effect size, power, and alpha)—$v = N - \text{numberOfPredictors} - 1$; thus, $N = v + \text{numberOfPredictors} + 1$:

```{r}
multipleRegressionSampleSizeModel <- pwr::pwr.f2.test(
  f2 = effectSize_fSquared,
  power = power,
  sig.level = alpha,
  u = numPredictors)

multipleRegressionSampleSizeModel

vNeeded <- multipleRegressionSampleSizeModel$v
sampleSizeNeeded <- vNeeded + numPredictors + 1
sampleSizeNeeded
```

```{r}
pwr::pwr.f2.test(
  power = power,
  sig.level = alpha,
  u = numPredictors,
  v = N - numPredictors - 1)
```
